---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r}
activity <- read.csv("activity.csv")
attach(activity)
summary(activity)
str(activity)
```


## What is mean total number of steps taken per day?

```{r}
#sum(steps, na.rm=TRUE)
stepsPerDay <- tapply(steps, date, sum)
hist(stepsPerDay, breaks=30)
```
The mean steps per day is:
```{r}
meanPerDay <- mean(stepsPerDay, na.rm=TRUE)
meanPerDay
```
The median steps per day is:
```{r}
medianPerDay <- median(stepsPerDay, na.rm=TRUE)
medianPerDay
#plot(meanPer5min ~ intervals, type="l")
```


## What is the average daily activity pattern?

Each day is divided into 5-minute intervals.  
```{r}
intervals <- factor(interval)
intervalsPerDay <- length(levels(intervals))
meanPer5min <- tapply(steps, intervals, mean, na.rm=TRUE)
plot(levels(intervals), meanPer5min, type="l")
```
```{r}
maxPer5min <- max(meanPer5min)
```
Max steps per 5-min interval: `r maxPer5min`
```{r}
max5min <- 0
for (i in 1:nrow(meanPer5min)) {
      if (meanPer5min[i] == maxPer5min) {
            max5min <- i
      }
}
```
The number of the interval, out of `r intervalsPerDay`, with the most steps averaged over all days is `r max5min`.


## Imputing missing values

Number of missing ("NA") intervals:
```{r}
stepsNA <- steps[steps=="NA"]
length(stepsNA)
```

```{r}
stepsImputed <- steps
stepsWeekday <- 0
stepsWeekendDay <- 0

stepsImputedNAi <- 0
for (i in 1:length(steps)) {
      if (is.na(steps[i])) {
            # Deal with steps that were all NA:
            if (is.na(meanPer5min[i])) {
                  meanPer5min[i] <- stepsImputed[i-1]
            }
            stepsImputed[i] <- meanPer5min[i]
      }
      if (is.na(stepsImputed[i]) && stepsImputedNAi == 0) {
            stepsImputedNAi <- i
      }
      if (!grepl('^S',weekdays(as.Date(date[i])))) {
            stepsWeekday <- stepsWeekday + stepsImputed[i]
      } 
      else {
            stepsWeekendDay <- stepsWeekendDay + stepsImputed[i]
      }
}
stepsImputedNAi
```

```{r}
stepsImputedPerDay <- tapply(stepsImputed, date, sum)
hist(stepsImputedPerDay, breaks=30)
```

```{r}
meanImputedPerDay <- mean(stepsImputedPerDay, na.rm=TRUE)
medianImputedPerDay <- median(stepsImputedPerDay, na.rm=TRUE)
```
The mean imputed steps per day is: `r meanImputedPerDay`

The median imputed steps per day is: `r medianImputedPerDay`


## Are there differences in activity patterns between weekdays and weekends?
```{r}
# Find numbers of weekday and weekend days in 'dates':
dates <- levels(date)
nWeekendDays <- 0
for (d in dates) {
      if (grepl('^S',weekdays(as.Date(d)))) {
            nWeekendDays <- nWeekendDays + 1
      }
}
nWeekdays <- length(dates) - nWeekendDays

#stepsWeekday
#nWeekdays
stepsPerWeekday <- stepsWeekday / nWeekdays

#stepsWeekendDay
#nWeekendDays
stepsPerWeekendDay <- stepsWeekendDay / nWeekendDays
```
There were `r stepsPerWeekday` average steps per weekday, and 
`r stepsPerWeekendDay` average steps per weekend day.
